---
title: "Olympics 2026"
description: |
  Analysis of 2026 Olympics data from TidyTuesday
author: Jamie Paik
date: September 28, 2026
format: 
  html:
    code-fold: true
    code-tools: true
    source: true
    code-overflow: wrap

execute: 
  warning: false
  message: false
---
```{r}
#| echo: false
library(tidyverse)
library(openintro)
```

This TidyTuesday dataset is about the 2026 Winter Olympics in Milan-Cortina, Italy. The dataset contains detailed information about all 1,866 Olympic events, including both competition and training sessions across various winter sport disciplines.

The data frame combines data from three sources:
Olympics.com - Official detailed daily schedule (events, times, venues)
Wikipedia - Competition schedule overview table
Kaggle - Venue information dataset

Credits for data source:
[Github Link](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-02-10/readme.md)

@chendaniely, Daniel Chen, curated the dataset


```{r}
#| echo: false
dataset_url = "https://raw.githubusercontent.com/chendaniely/olympics-2026/refs/heads/main/data/final/olympics/olympics_events.csv"
olympics <- read_csv(dataset_url) #reading csv data into Positron
```

```{r}
olympics_ski <- olympics |>
    filter(str_detect(discipline_name,"Skiing"))

#filtering dataset for sports that are related to skiing only
```

```{r}
olympics_ski <- olympics_ski |>
  mutate(medal = if_else(is_medal_event == TRUE, "Medal", "Non-medal"))

#mutating the is_medal_event variable to show actual description of medal game status
```

```{r}
# segmented bar chart
ggplot(olympics_ski, aes(x = discipline_name, fill = medal)) +
  geom_bar(position = "fill") +
  labs(
    x = "Type of Ski Sport",
    y = "Proportion of Medal Events",
    title = "Proportion of Medal Events for Olympic Ski Sports"
)
```

This bar chart illustrates the proportion of ski-related Olympic sports that were medal vs non-medal in the 2026 Winter Olympics. Both Alpine Skiing and Cross-Country Skiing are judged strictly by clock time, while Freestyle Skiing involves a mix of timed racing and subjective judge scoring. It is clear that Cross-Country Skiing has proportionately the most medal games, while Freestyle Skiing has the smallest proportion, which might be influenced by the fact that there is a degree of subjective judgement required, hence a team of qualified judges are needed for a medal game.

```{r}
# Create new variable for game duration in minutes

olympics <- olympics |>
  mutate(duration = as.numeric(end_time - start_time) / 60)

```

```{r}

# scatter plot
ggplot(olympics, aes(x = discipline_name, y = duration)) +
  geom_point() +
  labs(
    x = "Olympic Sport",
    y = "Duration of Game (mins)",
    title = "Duration of each game for each olympic sport") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The scatterplot illustrates the duration of each game (in minutes) for each 2026 Winter Olympic sport. Figure skating seems to have games with the longest game time at maximum 250 minutes (around 3 hours) and sports that are strictly judged by speed and time, such as Cross-Country Skiing, Speed Skating, or Short Track Speed Skating have the shortest games.